GravExplain: Continuous gravitational wave searches in a table-top experiment

### 2020-02-12 paper/paper_main.tex
75th day of Summer

@@

% overview
In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous gravitational waves to a general audience. 
The techniques used are very similar to those used to search for continuous waves in LIGO and Virgo data.
In Section~\ref{sec:ifo} we detail the interferometer set-up, Section~\ref{sec:single_tone} demonstrates observing a single tone, Section~\ref{sec:viterbi_wandering} details tracking a wandering frequency tone, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech. We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.

@@

Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal. Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied. \jam{Should I show photodiode Viterbi results?}

@@

@@

% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio. However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference. An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
% % Visually it is impossible to confirm this for high frequencies, but it remains possible.

@@

\jam{BATTERY POWER}

For further speech enhancement, the optimal filter would involve a hidden-Markov-model trained to English phonemes. Alternatively, machine learning solutions also exist through-out the field.
% SEGAN?

% any other wild speculation?

@@

@@

\section{Open-source code}
This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.

@@

\section{Guide to reproducing results}
\label{app:reproducing_results}
% as for a 3rd year undergraduate
\jam{Add this in, lots of advice and detail!}

>>>>

\jam{write this!}

@@

% overview
In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous gravitational waves to a general audience. 
The techniques used are very similar to those used to search for continuous waves in LIGO and Virgo data.
As an overview, Section~\ref{sec:ifo} details the interferometer, Section~\ref{sec:single_tone} describes observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech. We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.

@@

Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal. Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.

@@

\jam{Explain MMSE as in the second talk, be honest in some way}

@@

% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio. However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference. An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
% Visually it is impossible to confirm this for high frequencies, but it remains possible.

@@

\jam{BATTERY POWER}
% any other wild speculation?

@@

For further speech enhancement, the optimal filter would involve a hidden-Markov-model trained to English phonemes. Alternatively, machine learning solutions also exist through-out the field.

@@

\section{Open-source code}
\label{app:code}
This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.

@@

\section{Additional guide to reproducing results}
\label{app:reproducing_results}
% as for a 3rd year undergraduate

Besides the directions given in the rest of this article, there are some observations and technical details useful to someone wanting to reproduce the results shown herein. For the analysis and signal processing performed the best reference is Appendix~\ref{app:code} for the open-source code.

For the experiment, since a Michelson interferometer is such a standard configuration, we’ll not detail it further here, see any optics manual for more information. To obtain an independent speaker cone we recommend carefully dismantling a budget commercial speaker system to get at the speaker within. To attach the speaker to one of the mirrors, we wedged it into part of the mirror mount and used commercial adhesive putty to keep it in place. If the other speaker in the pair remains, keep it as far away from the experiment and as damped as possible else it may act as a second, interfering source.
% we used a Class 2 LDS5 ThorLabs laser at 532nm

For the photodiode circuit, we used a Raspberry Pi and an MCP3008 ADC to monitor the analog signal. This could also be done with the analog pins on an Arduino. Be careful of the gain through the LM358 op-amps in the circuit though, else the ADC in either case could threshold the signal. Also be ready to change the values in the Sallen-Key filter to match your own sampling rate. The photodiode was an OSRAM BPW21, we mounted it through a discarded piece of speaker mesh (similar to black cloth over a frame) and placed another piece over the face to not over expose and threshold the signal. For the rest of the circuit, we used standard components assembled as in \ref{app:circuit_diagram}.

\jam{Add this in, lots of advice and detail!}
